---
title: "Module 2: Import and Structure"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    navigation-mode: vertical
    footer: "wearabler"
slide-level: 2
---

# Main Question

How are time and axes represented?

# Teaching Dataset

The raw `.gt3x` and `.cwa` files in `data/`.

## Core Representation

Raw accelerometry is, at its core, a time series with four essential columns:

| Column | Meaning |
|---|---|
| `time` | A datetime timestamp identifying when the sample was recorded |
| `x` | Acceleration along the device x axis |
| `y` | Acceleration along the device y axis |
| `z` | Acceleration along the device z axis |

The timestamp is not merely a row number. Its timezone, precision, regularity, and gaps determine day boundaries, sleep windows, wear-time intervals, and the validity of all downstream summaries. The x, y, and z columns are the core of raw accelerometry; device headers and processing history are essential accompanying metadata.

## Read the Local Files with `actiread`

`actiread` provides a common R-facing import layer for the two file types used in this course. Import first, inspect second, and standardize before calculating a metric.

```{r}
#| eval: false
library(actibase)
library(actiread)

gt3x_file <- "data/TAS1H30182791 (2026-04-29).gt3x"
cwa_file <- "data/114890_0000000000.cwa"

gt3x_raw <- acti_read_gt3x(gt3x_file, verbose = FALSE)
cwa_raw <- acti_read_cwa(cwa_file, verbose = FALSE)

gt3x <- acti_standardize_data(gt3x_raw)
cwa <- acti_standardize_data(cwa_raw)

head(gt3x[, c("time", "x", "y", "z")])
attributes(gt3x)
```

Before proceeding, verify the timezone, sampling frequency, acceleration units, start and end times, duplicate timestamps, and gaps. Preserve the original file and keep import metadata with the derived table.

## Read the Axivity File with `GGIRread`

`GGIRread` is a useful independent reader for Axivity `.cwa` data. Its output needs to be converted to the course's common `time`, `x`, `y`, `z` convention before it can be compared with another reader.

```{r}
#| eval: false
ax3 <- GGIRread::readAxivity(
  filename = "data/114890_0000000000.cwa",
  desiredtz = "America/New_York"
)

str(ax3)
# Extract the time and x/y/z components, then rename them to the common schema.
```

## Read the Axivity File with Python `actipy`

`reticulate` lets an R analysis call Python's `actipy` without leaving the project. This is especially helpful for comparing reader and preprocessing implementations. Set processing options explicitly; do not silently accept defaults as a study protocol.

```{r}
#| eval: false
library(reticulate)

actipy <- import("actipy")
result <- actipy$read_device(
  "data/114890_0000000000.cwa",
  lowpass_hz = 20,
  calibrate_gravity = TRUE,
  detect_nonwear = FALSE,
  resample_hz = 50
)

ax3_py <- py_to_r(result[[1]])
ax3_info <- py_to_r(result[[2]])
head(ax3_py)
ax3_info[c("SampleRate", "CalibErrorBefore(mg)", "CalibErrorAfter(mg)")]
```

## Learning Check

Can you state the sampling frequency and timezone for each imported file, and can you create a four-column table named `time`, `x`, `y`, and `z` without changing the raw values? If not, the import step is not complete.
