---
title: "Module 2: Reading and Processing Data"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    footer: "Wearable-R: https://github.com/jhuwit/wearabler"
    logo: ../assets/jhu.png
    css: ../assets/logo.css
bibliography: ../assets/references.bib    
slide-level: 2
---


```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(targets)
library(activerse)
options(digits.secs = 3)
knitr::opts_chunk$set(echo = FALSE)
```

# Materials

Materials are at: https://github.com/jhuwit/wearabler

Can use `git clone` or download via: https://github.com/jhuwit/wearabler/archive/refs/heads/main.zip

# Loading `targets` and environment

Using [`renv`](https://rstudio.github.io/renv/articles/renv.html) allows us to specify the packages used to build this pipeline:

```{r}
#| eval: false
renv::restore()
```

The [`targets`](https://books.ropensci.org/targets/) pipeline located [in the GitHub repo](https://github.com/jhuwit/wearabler/blob/main/_targets.R) will do the processing (and allows us to cache the pipeline).

```{r}
#| eval: false
targets::tar_make()
```

# Data: Personal Recording (from John)

- `TAS1H30182791 (2026-04-29).gt3x` - wore GT9X device
- `114890_0000000000.cwa` - wore AX3 device
- `AI15_MOS2D09170398_2017-10-30.gt3x` - 7 days of data from  [[@chadwell2019upper]](https://springernature.figshare.com/articles/dataset/Unprocessed_raw_30Hz_acceleration_data_stored_as_gt3x/7946189)

# Pipeline Overview

```{r}
#| eval: false
# Read in Raw data
data_gt3x = acti_read_gt3x(file_gt3x)
# Make minute-level data
counts_and_wear = acti_process(data_gt3x, method = "choi")
# get day-level inclusion
actibase::create_day_inclusion(counts_and_wear, min_required = 1296L)
counts_and_wear = add_day_inclusion(counts_and_wear, min_required = 1368L)
# Calculate daily measures
daily = counts_and_wear |> 
  filter(is_included) |> 
  group_by(date) |> 
  summarise(
    total_counts = sum(counts, na.rm = TRUE),
    total_counts_log10 = sum(counts_log10, na.rm = TRUE)
  ) |> 
  ungroup()
# do testing after inclusion criteria applied
daily |> 
  # usually group_by(id)
  mutate(n_days = n_distinct(date)) |> 
  filter(n_days >= 3) |> 
  summarise(
    total_counts = mean(total_counts),
    total_counts_log10 = mean(total_counts_log10)
  )
```


# Pipeline Overview

Read in Data (also `acti_read_cwa`):
```{r}
#| eval: false
#| echo: true
data_gt3x = acti_read_gt3x(file_gt3x)
```

Create minute level summaries, including wear inclusion criteria

```{r}
#| eval: false
#| echo: true
counts_and_wear = acti_process(data_gt3x, method = "choi")
counts_and_wear = add_day_inclusion(counts_and_wear, min_required = 1368L)
```

# Pipeline Overview

Keep days with minimum wear time, calculate day-level summaries
```{r}
#| eval: false
#| echo: true
daily = counts_and_wear |> 
  filter(is_included) |> 
  group_by(date) |> 
  summarise(
    total_counts = sum(counts, na.rm = TRUE),
    total_counts_log10 = sum(counts_log10, na.rm = TRUE)
  ) |> 
  ungroup()
```

# Pipeline Overview

Keep participants with minimum number of days observed, create "average" day activity:
```{r}
#| eval: false
#| echo: true
# all_daily = dplyr::bind_rows(daily_1, daily_2, ...)
daily |> 
  # usually group_by(id)
  mutate(n_days = n_distinct(date)) |> 
  filter(n_days >= 3) |> 
  summarise(
    total_counts = mean(total_counts),
    total_counts_log10 = mean(total_counts_log10)
  )
```

# Pipeline Overview

Statistics!


# Reading in the Data

```{r}
#| eval: false
#| echo: true
library(actiread)
file_gt3x = here::here("data/TAS1H30182791 (2026-04-29).gt3x")
file_cwa = here::here("data/114890_0000000000.cwa")

data_gt3x = actiread::acti_read_gt3x(file_gt3x)
data_cwa = actiread::acti_read_cwa(file_cwa)
head(data_gt3x, n = 3)
head(data_cwa, n = 3)
```


```{r load_raw_data}
#| echo: false
tar_load(data_cwa)
tar_load(data_gt3x)
head(data_gt3x, n = 3)
head(data_cwa, n = 3)
```


# Reading CWA files

Different implementations of reading in the data
* we recommend `acti_py_read_cwa` (closer to OMGUI)


```{r}
#| eval: false
#| echo: true
# uses actipy implementation not `GGIR:readAxivity`
data_cwa_py = actiread::acti_py_read_cwa(file_cwa)
head(data_cwa, n = 3)
head(data_cwa_py, n = 3)
```

```{r load_raw_data_py}
#| echo: false
tar_load(data_cwa_py)
head(data_cwa, n = 3)
head(data_cwa_py, n = 3)
```

# Summary

- Use `actiread::acti_py_read_cwa` for CWA
- Use `actiread::acti_read_gt3x` for GT3X
- We will show small and large example GT3X data


# Core Representation

Core raw accelerometry is time series 3 axes:

| Column | Meaning |
|---|---|
| `time` | A datetime timestamp identifying when the sample was recorded |
| `x`,`X` | Acceleration along the device x axis |
| `y`,`Y` | Acceleration along the device y axis |
| `z`,`Z` | Acceleration along the device z axis |



# Data Attributes {.small}

- Sample rate - sampling frequency in Hz: `get_sample_rate`
- Dynamic accelerometry range in `g`: `get_dynamic_range`

```{r attrs}
#| echo: true
get_sample_rate(data_gt3x)
attr(data_gt3x, "sample_rate")
get_sample_rate(data_cwa)

get_dynamic_range(data_gt3x)
get_dynamic_range(data_cwa)

names(attributes(data_gt3x))
```

# Data Header from GT9X Device

```{r}
#| echo: true
attr(data_gt3x, "header")
```
# Data Header from AX3 Device

```{r attr_header}
#| echo: true
attr(data_cwa, "header")
```



# Timestamps Need Their Own QC {.small}

* Timestamp: use its timezone, precision, regularity, and gaps determine day boundaries, sleep windows, wear-time intervals, and the validity of all downstream summaries. 

* Sub-second algorithms depend on fractional seconds. Display enough precision to see them (for example, `options(digits.secs = 3)`) and check observed intervals rather than assuming that every nominal 80 Hz file is exactly 80 Hz at every timestamp. 

* Rounding milliseconds during a CSV export can corrupt joins, inferred sampling rates, resampling, and step algorithms.

# Timestamps Need Their Own QC {.small}

* Use an explicit timezone. 
* UTC avoids daylight-saving discontinuities during processing
* Local time is often needed for sleep windows, diary linkage, and calendar-day summaries. 
* Record the original timezone, the analysis timezone, and every conversion between them.

* Verify the timezone, sampling frequency, acceleration units, start and end times, duplicate timestamps, and gaps. Preserve the original file and keep import metadata with the derived table.

```{r any_dup}
#| echo: true
anyDuplicated(data_gt3x$time)
anyDuplicated(data_cwa$time)
```



# Plot raw data

- `acti_tidy_axes` will reshape the time/XYZ data to a long format for plotting

```{r tidy_axes}
#| echo: true
long = data_gt3x |> acti_tidy_axes()
head(long, n = 3)
long |> count(axis)
```

# Plot raw data

```{r tidy_axes_plot}
#| echo: true
long |> 
  ggplot(aes(x = time, y = value, colour = axis)) + geom_step()
```

# If you plot less with more data, something is going in the wrong direction <br>- Karl Broman

# Plotting raw data - big

But the if the data is > 7 days with 80Hz,

7 days * 24 hours/day * 60 minutes/hour * 60 seconds/minute * 80 samples/second
```{r}
7*24*60*60*80
```
A lot of Points!

# Plotting

* Many times plot at a minute level
  * Get time limits for checking
  * Zoom in raw data at certain intervals

QC: Check for odd things

- Zeroes 
- Flat data
- Really high values


# Idle sleep mode - 0s/missing

* Idle sleep mode - battery saving option when enabling and setting up ActiGraph devices.  
* Stops recording after no movement after while

* It's a hassle, don't use it unless you have to.  

* It causes "gaps" in the GT3X file that are truly missing, but this is an issue with many processing methods (gaps), so you can "impute" zeroes into the data set, which aren't correct either. 


# How long is not wearing/zero

Question:

* How long is low/no activity actually someone not wearing the device?

# Non-wear Estimation

In a nutshell: 

1. define threshold for no activity
2. define time threshold for how long it lasts (e.g. 90 minutes)
3. do you allow for any "spikes"?

# Non-wear Estimation

- [@van2011estimation]: Non-wear is a 30-min non-overlapping raw-acceleration window with SD <3 mg or range <50 mg in at least 2 of 3 axes.
- [@van2013separating]: now 60-min rolling window w/15-min steps, 00:00–01:00, then 00:15–01:15
- **[@CHOI2011]**: Classifies ≥90 minutes of zero activity counts, allowing brief interruptions, as non-wear. 
- [@troiano2008physical]: Defines non-wear as ≥60 consecutive zero-count minutes, permitting short low-count interruptions.


# What about GGIR

* Very popular
* Has changed/added a number of algorithms over the years
  * Good: state of the art, bad: some reproducibility concerns over time
* Want something more modular
  * GGIR creates a "buffet" of activity measures
* A lot of python modules have made considerable contributions
* Used in underlying functions, but not main GGIR runner

# Gravity Correction/Calibration 

Gravity correction [@gravity] or gravity calibration: try to correct data that is miscalibrated
- estimates scale and shift parameters fitting to unit sphere

**Note: Not all pipelines employ this, including a number of pipelines for ActiGraph data.**

# Gravity Calibration: Steps

1. Read the original tri-axial `X`, `Y`, and `Z` acceleration; do not use imputed zeroes.
2. Find short stationary windows (for example, 10 seconds with low SD on all three axes).
3. Use the stationary vectors to estimate each axis's offset and scale so their vector magnitude lies on the unit sphere.
4. Apply the estimated offset and scale to every raw sample, then calculate metrics such as ENMO or activity counts from the corrected signal.
5. Record the calibration parameters and inspect the before/after gravity error; flag a file when there are too few eligible stationary windows.

# Gravity Calibration: Steps

- `GGIR::g.calibrate()`
- `agcounts::agcalibrate`
- `acti_calibrate` - uses `agcounts::agcalibrate`


```{r g_calibrate}
targets::tar_load(data)
```

```{r calibrate_cal_estimates_load}
#| echo: true
#| results: hide
#| cache: true
data_calibrated = acti_calibrate(data)
```


# Gravity Calibration

```{r calibrate_cal_estimates}
calibration = attr(data_calibrated, "calibration")
calibration$scale
calibration$offset
head(data)
head(data_calibrated)
```

# No calibration done

- Likely should be used in analysis
- Most ActiGraph pipelines do not use it
- Can be sensitivity analysis

# Activity Counts

* Choi/Troiano use (ActiGraph) Activity Counts

* [@neishabouri2022quantification] - made open source the activity count calculation
* [`agcounts`](https://github.com/bhelsel/agcounts) R package

```{r ac_calculation}
#| echo: true
counts_gt3x = acti_calculate_counts(data_gt3x, verbose = FALSE)
counts_cwa = acti_calculate_counts(data_cwa)
```


# Activity Counts

```{r ac_plot}
#| echo: true
counts_gt3x |> 
  ggplot(aes(x = time, y = counts)) + geom_step()
counts_cwa |> 
  ggplot(aes(x = time, y = counts)) + geom_step()
```

# Estimate Non-wear

* Calculate wear using `acti_calculate_wear`, uses Choi by default

```{r}
#| echo: true
wear_gt3x = acti_calculate_wear(counts_gt3x)
head(wear_gt3x, n = 3)
wear_cwa = acti_calculate_wear(counts_cwa)
head(wear_cwa, n = 3)
```

# Just use `acti_process`

```{r show_load_minute_gt3x}
#| echo: true
#| eval: false
minute_gt3x = acti_process(data_gt3x, verbose = FALSE)
minute_cwa = acti_process(data_cwa, verbose = FALSE)
```

```{r load_minute_gt3x}
#| echo: false
minute_gt3x = tar_read(data_gt3x_proc)
minute_cwa = tar_read(data_cwa_proc)
```

- Creates counts, non-wear, and joins them
- Simple starting point for analysis
- Will cover other measures

# Plotting

<!-- data_gt3x |> filter(time < ymd_hms("2026-04-29 11:39:00.000")) |> select(X, Y, Z) |> as.matrix() |> matrixStats::colVars() -->

```{r running_minute}
#| echo: true
minute_gt3x |> 
  ggplot(aes(x = time, y = counts)) + 
  geom_segment(
    aes(x = time, xend = time,
        y = -Inf, yend = Inf, color = wear), alpha = 0.25) +
  geom_step() 
```

# Reading Full Data

<!-- data_gt3x |> filter(time < ymd_hms("2026-04-29 11:39:00.000")) |> select(X, Y, Z) |> as.matrix() |> matrixStats::colVars() -->

We can read in the data from [@chadwell2019upper] and run the minute level data with wear/non-wear:

```{r load_minute_show}
#| eval: false
#| echo: true
data = acti_read_gt3x(here::here("data/AI15_MOS2D09170398_2017-10-30.gt3x"))
minute = acti_process(data)
```

```{r load_minute}
#| echo: false
minute = tar_read(data_gt3x_proc)
```

# Plotting Full Data

```{r full_minute_plot}
#| cache: true
#| echo: true
minute |> 
  ggplot(aes(x = time, y = counts)) + 
  geom_segment(
    aes(x = time, xend = time,
        y = -Inf, yend = Inf, color = wear), alpha = 0.25) +
  geom_step() 
```

# Plotting Day-level


```{r plot_full_data_show}
#| eval: true
#| echo: false
#| cache: true
tar_load(data_proc)
data_proc |> 
  mutate(date = lubridate::as_date(time),
         hourtime = hms::as_hms(time)) |> 
  ggplot(aes(x = hourtime, y = counts)) + 
  geom_segment(
    aes(x = hourtime, xend = hourtime,
        y = -Inf, yend = Inf, color = wear), alpha = 0.25) +
  geom_step() +
  facet_wrap(~ date, ncol = 1)
```

# Day Level Inclusion

Usually determined by number of minutes wear time

- 1296 - 90% of 1440 minutes [@schrack2014assessing]
- 1368 - 95% of 1440 minutes [@leroux2024nhanes]
- 10 hours of wear (may be contiguous or not) [@troiano2008physical]
- 10 hours daytime wear (6AM-12AM, not good for sleep) [@dillon2018associations]


# Day Level Inclusion

* `create_day_inclusion` - creates a table for inclusion

```{r create_day_inclusion}
#| echo: true
minute |> 
  actibase::create_day_inclusion(min_required = 1296L) |> knitr::kable()
```

# Day Level Inclusion

* `add_day_inclusion` - creates and joins with time-series

```{r day_inclusion}
#| echo: true
minute |> 
  actibase::add_day_inclusion(min_required = 1296L) |> 
  select(time, counts, wear, n_minutes_wear, is_included) |> knitr::kable()
```



# Summary

- Raw data to minute level counts
- Non-wear estimated from low activity for long periods of time
- `acti_process` can do that and merge the data
- Plotting the data typically using `geom_step` and color by `geom_segment`
- Day level inclusion usually based on wear


# References
