---
title: "Module 3: Alternative Metrics"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    footer: "Wearable-R: https://github.com/jhuwit/wearabler"
    logo: ../assets/jhu.png
    css: ../assets/logo.css
bibliography: ../assets/references.bib    
slide-level: 2
---


```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(targets)
library(activerse)
options(digits.secs = 3)
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
tar_load(data_gt3x)
```


```{r}
#| eval: true
#| echo: false
#| results: hide
#| cache: true
measures = acti_calculate_measures(data_gt3x, flag_data = FALSE)
```


# Alternative Metrics: AI

@bai2016activity introduced the Activity Index (AI), an open-source metric that operates on raw data, has units in SI units ($g$ or milli-$g$) and has more sensitive ROC curves for moderate and vigorous physical activities compared to AC and was shown to be more sensitive to sedentary and light activities than ENMO.

For a given window $H$, AI is calculated for a given participant is calculated using $AI_{0}$ (no baseline variance estimated as in paper):

$$
\text{AI}(t; H)_0 = \sqrt{ \frac{1}{3} \sum\limits_{m=1}^3 \sigma_{m}^2(t; H)  }
$$


which is the square root of the average variance of the 3 axes, and provides an average standard deviation. 

@bai2016activity indicated that the window to calculate AI was 1 second and "any aggregated AI was obtained by summing up all the adjacent 1-second AIs within that period".  Thus, to calculate AI at a minute level, we would calculate it at a second level and then add the 60 $\text{AI}_0$ values together.

# Alternative Metrics: AI

```{r}
#| echo: false
measures |> 
  ggplot(aes(x = time, y = AI)) + geom_step()
```



# Alternative Metrics: Mean Amplitude Deviation (MAD)

@mad recommended the Mean Amplitude Deviation (MAD) as a measure that was an open-source, in SI units, and had good discriminating power for separating activity levels from each other.  The MAD is calculated as the mean of the absolute difference between the acceleration and the mean acceleration for a given epoch.

For a given epoch of length $m$, and where $\text{VM}$ represents the vector magnitude of the signal, the calculation of MAD is:
$$
\text{MAD} = \frac{1}{m} \sum_{i=1}^{m} \left|\text{VM}_i- \overline{\text{VM}}_m \right|
$$
For a given epoch, @mad recommended taking the mean value of MAD for summarization and recommended 5-second epochs.

# Alternative Metrics: MAD

```{r}
#| echo: false
measures |> 
  ggplot(aes(x = time, y = MAD)) + geom_step()
```

# Alternative Metrics: Monitor-Independent Movement Summary (MIMS) Units

@mims introduced the Monitor-Independent Movement Summary (MIMS) unit that was to be an "open-source, universal acceleration summary metric that accounts for discrepancies in raw data among research and consumer devices". 

* raw data interpolated of the signal to 100H 
* extrapolates signal for regions that have hit the maximum/minimum acceleration units 
* band-pass filters from 0.2 to 5Hz,
* absolute value of the area under the curve (AUC) using a trapezoidal rule
* truncates low signal values to $0$.

@karas2022comparison found a correlation of $\geq 0.97$ between AC and MIMS units.  

```{r}
#| echo: false
measures |> 
  ggplot(aes(x = time, y = MIMS)) + geom_step()
```

# Alternative Metrics: ENMO

* Euclidean Norm Minus One (ENMO)
* At raw level $\sqrt(X^2+Y^2+Z^2) - 1$.  Usually truncated at 0.
* Mean values (not sums) are calculated for different windows

```{r}
#| echo: false
measures |> 
  ggplot(aes(x = time, y = ENMO_t)) + geom_step()
```


# Calculate Them All

* The `acti_calculate_measures` will run AC, AI, MIMS, and MAD
```{r}
#| eval: false
#| cache: true
measures = acti_calculate_measures(data_gt3x)
```


