---
title: "Module 3: Import and Structure"
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    footer: "Wearable-R: https://github.com/jhuwit/wearabler"
    logo: ../assets/jhu.png
    css: ../assets/logo.css
bibliography: ../assets/references.bib    
slide-level: 2
---


```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(targets)
library(activerse)
options(digits.secs = 3)
knitr::opts_chunk$set(echo = FALSE)
tar_load(data_gt3x)
```

# Calculate Them All

* The `acti_calculate_measures` will run AC, AI, MIMS, and MAD
```{r}
#| eval: false
#| cache: true
measures = acti_calculate_measures(data_gt3x)
```



# Do plots of 10 people

# Lasagna plot of NHANES


## Other Walking Measures


## Step Counts

Open source method for estimating stepcounts: https://github.com/OxWearables/stepcount [@stepcount], ported using `reticulate`. 

Must load module and then run (can use GPU):

```{r code_stepcount}
#| eval: false
actimetrics::py_acti_calculate_stepcount()
```

```{r get_stepcount}
#| echo: false
sc = readr::read_rds(here::here("data/stepcount.rds"))
```

```{r}

```

If no model is specified, the model is downloaded to a temporary directory, but users can use the `stepcount::sc_download_model` function to download the model file, store it, and pass it via the `model_path` argument. 

## Step Counts

:::: {.columns}

::: {.column width="50%"}

The output of `stepcount` is a list of different pieces of information, including information about the device:

```{r stepcount_output}
names(sc)
```

A `data.frame` of flags for walking:

```{r stepcount_walking}
head(sc$walking)
```


:::

::: {.column width="50%"}

The main output of interest: `data.frame` of estimated steps:

```{r stepcount_steps}
head(sc$steps)
```

```{r plot_steps}
#| echo: false
sc$steps |> 
  ggplot(aes(x = steps)) + geom_histogram() + xlab("Number of Steps in 10s")
```

:::
::::

## Step Counts Aggregation

The `stepcount` method estimates steps in 10-second windows, which we can aggregate into minutes:

```{r sc_steps_minute}
sc60 = sc$steps |> 
  mutate(time = floor_date(time, "1 minute")) |> 
  group_by(time) |> 
  summarise(steps = sum(steps), .groups = "drop")
head(sc60)
```

## Steps Per Minute


```{r plot_sc60_show}
#| eval: false
sc60 |> 
  mutate(date = lubridate::as_date(time),
         hourtime = hms::as_hms(time)) |> 
  ggplot(aes(x = hourtime, y = steps)) + 
  geom_step() +
  facet_wrap(~ date, ncol = 1)
```


```{r plot_sc60}
#| dependson: plot_sc60_show
#| echo: false
#| out-width: "100%"
<<plot_sc60_show>>
```
********


# Step Count

# Step Count Plot

# Actinet

# Actinet Plot


# `actipy`

# `pygt3x`

# Pfizer Package


# MVPA definition

- Montoye thresholds
- bring in actinet vs. AC discussion


# Overall premise (is all activity created equal)
## Fragmentation metrics
- arc tools

## Weekend warrior
### first defined
### WHy - all created equal

### Fragmentation pattern/cronotype/etc

# Mean analysis (discuss survey weights and sample weights (non-wear))
# FPCA (discuss survey weights)
- Recommend the book

# FUI
# CV AUC for mortality

# 




# Step Count Plot


# Threshold for Inclusion
- 1296
- 1368
- 10 hours
- 10 hours daytime

# MPVA/Sedentary/LPA 
- threshold
- threshold for measures (Montoye)

# MPVA/Sedentary/LPA history




```{r}
#| eval: false
raw |>
  dplyr::summarise(
    duplicate_times = sum(duplicated(time)),
    smallest_interval_s = min(diff(as.numeric(time))),
    largest_interval_s = max(diff(as.numeric(time)))
  )
```



```{r}
# actiquantiles::
```

