---
title: "Module 3: Import and Structure"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    footer: "Wearable-R: https://github.com/jhuwit/wearabler"
    logo: ../assets/jhu.png
    css: ../assets/logo.css
bibliography: ../assets/references.bib    
slide-level: 2
---


```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(targets)
library(activerse)
options(digits.secs = 3)
knitr::opts_chunk$set(echo = FALSE)
tar_load(data_gt3x)
```

# Calculate Them All

* The `acti_calculate_measures` will run AC, AI, MIMS, and MAD
```{r}
#| eval: false
#| cache: true
measures = acti_calculate_measures(data_gt3x)
```





# Overall premise (is all activity created equal)
## Fragmentation metrics
- arc tools

## Weekend warrior
### first defined
### WHy - all created equal

### Fragmentation pattern/cronotype/etc

# Mean analysis (discuss survey weights and sample weights (non-wear))
# FPCA (discuss survey weights)
- Recommend the book

# FUI
# CV AUC for mortality

# 



# Do plots of 10 people

# Lasagna plot of NHANES


## Other Walking Measures



# Step Count Plot






```{r}
#| eval: false
raw |>
  dplyr::summarise(
    duplicate_times = sum(duplicated(time)),
    smallest_interval_s = min(diff(as.numeric(time))),
    largest_interval_s = max(diff(as.numeric(time)))
  )
```



```{r}
# actiquantiles::
```
