---
title: "Module 7: Daily and Participant-Level Summaries"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    navigation-mode: vertical
    footer: "wearabler"
slide-level: 2
---

# Main Question

How do minute-level data become one interpretable measure per person?

# Teaching Dataset

Valid participant-days from a multi-participant study.

## Daily Summaries

For each valid day, summarize the selected metric over eligible wear minutes. Examples include mean ENMO, total minutes above a threshold, MIMS per day, count-based activity volume, sleep duration, and step count. A day-level summary must retain its denominator: number of eligible minutes or hours of wear.

```{r}
#| eval: false
daily <- minute_data |>
  dplyr::filter(wear, valid_day) |>
  dplyr::group_by(participant_id, date) |>
  dplyr::summarise(
    mean_enmo_mg = mean(enmo_mg),
    active_minutes = sum(enmo_mg >= 100),
    wear_minutes = dplyr::n(),
    .groups = "drop"
  )
```

## One Value Per Participant

A common participant-level outcome is the unweighted arithmetic mean of that person's valid daily summaries. It represents an estimate of the person's average day among observed valid days. Each retained day contributes equally; a day with more wear does not receive more weight simply because it contains more observed minutes.

```{r}
#| eval: false
participant <- daily |>
  dplyr::group_by(participant_id) |>
  dplyr::summarise(
    mean_daily_enmo_mg = mean(mean_enmo_mg),
    valid_days = dplyr::n(),
    .groups = "drop"
  )
```

This estimand intentionally does **not** account for the number of valid days in the mean itself. The number of days remains an important precision and QC variable and should be reported, used in sensitivity analyses, or incorporated into a different estimand when warranted.

## Learning Check

State whether your final value estimates an average minute, an average observed day, or an average participant. These are different quantities.
